
---
title: "Exploratory Data Analysis"
output:
  html_notebook:
    toc: yes
editor_options: 
  chunk_output_type: inline
---

Load tidyverse package and set plot theme
```{r, message = F}
library(tidyverse)

theme_set(theme_classic(base_size = 16))
```

Let's set up the discrimination example to see what a random sample would look like
```{r one_sample}
# 48 total employees, 24 Male and 24 Female
employees <- c(rep("Male", 24), rep("Female", 24))

# Let's randomly promote 35 employees (independent of gender)
promoted <- sample(employees, 35) 

# Compute the proportion of men and women promoted and not promoted
promoted_men <- sum(promoted == "Male")
promoted_women <- sum(promoted == "Female")
not_promoted_men <- 24 - promoted_men
not_promoted_women <- 24 - promoted_women

# Make a dataframe to show the results
data_frame(gender = c("Male", "Female", "Total"), 
           promoted = c(promoted_men, promoted_women, 
                        sum(promoted_men, promoted_women)),
           not_promoted = c(not_promoted_men, not_promoted_women,  
                            sum(not_promoted_men, not_promoted_women)),
           total = c(24, 24, 48))
```

What if we draw a lot of random samples?
```{r sample_function}
#Write a function to simulate 1 draw 
promoted_diff <- function() {
  
  # 48 total employees, 24 Male and 24 Female
  employees <- c(rep("Male", 24), rep("Female", 24))
  
  # Let's randomly promote 35 employees (independent of gender)
  promoted <- sample(employees, 35) 
  
  # Compute the difference in promoted men and women
  promoted_male <- sum(promoted == "Male")/24
  promoted_female <- sum(promoted == "Female")/24
  
  return(promoted_male - promoted_female)
}

# One function call
promoted_diff()
```

Sample and plot
```{r thousand_samples}
thousand_diffs <- data_frame(diff = replicate(1000, promoted_diff()))

ggplot(thousand_diffs, aes(x = diff)) + 
  geom_histogram(bins = 50)
```

Sample and plot along with our original result
```{r}
#find difference in the actual data
actual_diff <- (21/24) - (14/24)

thousand_diffs <- data_frame(diff = replicate(1000, promoted_diff()))

#add a vertical line to the random samples to show the empirical data
ggplot(thousand_diffs, aes(x = diff)) + 
  geom_histogram(bins = 50) + 
  geom_vline(xintercept = actual_diff, color = "darkred", size = 2) 
```

Where does our difference lie?
```{r}
# Compute the proportion of samples that the actual difference was greater than
mean(actual_diff >= thousand_diffs)
```

Our empirical difference in promotion probability is as big or bigger than almost all of the random samples we took! So probably something funny is going on, because this would be a really unusual result to happen by random chance.