
---
title: "Lab 5 - Foundations for statistical inference - Confidence intervals"
author: "Your name"
date: "Date of lab session"
output: 
  html_document: 
    highlight: tango
    theme: spacelab
---

```{r setup, include=FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

* * *

## Lab report

**Load data**
```{r load_data}
ames <- read_csv("https://dyurovsky.github.io/85309/data/lab5/ames.csv")
```

```{r set_seed}
set.seed(85309)
```

#### Exercise 1: 
```{r ex1}
ggplot(ames, aes(x = area)) + 
  geom_histogram(binwidth = 250)

ames %>%
  summarise(pop_med = median(area), 
            mu = mean(area),
            pop_iqr = IQR(area))  # third quartile, 75th percentile
```

The distribution is unimodal and slightly right-skewed. Because it is non-normal, we can describe it with robust statistics like median and iqr (above).

#### Exercise 2:
```{r ex2}
samp1 <- ames %>%
  sample_n(50)

ggplot(samp1, aes(x = area)) + 
  geom_histogram(binwidth = 250)

```

Compared to the population distribution, the sample distribution has a similar skewed shape but is noisier. The tails look less smooth.

#### Exercise 3:
```{r ex3}
samp2 <- ames %>%
  sample_n(50)

ggplot(samp2, aes(x = area)) + 
  geom_histogram(binwidth = 250)
```

Sample 2 has a lot of the same features as sample 1. It is right-skewed but not too much and unimodal like the population distribution. If we took a sample of size 100 and a sample of size 1000, the sample of size 1000 will look more like the population.

#### Exercise 4:
```{r ex4}
sample_50 <- function() {
  ames %>%
    sample_n(50, replace = TRUE) %>%
    summarise(x_bar = mean(area)) %>%
    pull()
}

sample_means50 <- tibble(sample = 1:1000,
                         mean = replicate(1000, sample_50()))

ggplot(sample_means50, aes(x = mean)) + 
  geom_histogram()


sampling_distribution_mean <- sample_means50 %>%
  summarise(mean = mean(mean)) %>%
  pull()
```

The sampling distribution looks approximately normal. It has a mean of `r sampling_distribution_mean`. This is very close to the population mean of 1500. There are 1000 elements of sample_means50 each of which is the sample mean of a sample of size 50.

#### Exercise 5:
```{r ex5}
sample_10 <- function() {
  ames %>%
    sample_n(10, replace = TRUE) %>%
    summarise(x_bar = mean(area)) %>%
    pull()
}

sample_means_small <- tibble(sample = 1:25,
                         mean = replicate(25, sample_10()))

ggplot(sample_means_small, aes(x = mean)) + 
  geom_histogram(binwidth = 100)
```

#### Exercise 6:

```{r ex6}
# enter your code for Exercise 6 here
```

### More Practice:

#### Exercise 7:

```{r ex7}
# enter your code for Exercise 7 here
```

#### Exercise 8:


```{r ex8}
# enter your code for Exercise 8 here
```

#### Exercise 9:

```{r ex9}
# enter your code for Exercise 9 here
```

#### Exercise 10:

```{r ex10}
# enter your code for Exercise 10 here
```
