---
title: "A brief intro to R"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r load_tidyverse}
library(tidyverse)
```

R can do math!
```{r}
4 + 3

2*6

20/4

3^2
```

You can store a certain number in a variable name

```{r}
x <- 5

x

x + 2

x + 1

x * 6

```

After you do math with your variable, you can store the answer in a new variable name
```{r}
a <- x + 5

a

b <- a * 3

b
```


If you store a new number in the variable name, it erases the old number and takes it’s place
```{r}
b <- 3

b

b <- b + 1

b
```


The name of your variable doesn’t have to just be a single letter. It can be any string of characters you want. (It has to start with a letter and can’t contain spaces. Some other characters also not allowed.)
```{r}
Pizza_Slice <- 2

4 * Pizza_Slice

cs <- 3

ps <- 2.5

ss <- 2.75

2*cs + ps
```


NOTE: Variable names are case sensitive!
```{r}  
ps

PS
```

You can store a whole list of numbers (a “vector”) in a single variable name. Easiest to think of this as a column of numbers in a spreadsheet and of the variable name as the heading of the column.
```{r}
x <- c(1,2,3)

x
```

You can do math with vectors in R
```{r}
x + 5
```


Again, you can store the answer in a new variable name if you’d like
```{r}
My_Answer <- x + 3

My_Answer
```


Here’s some more math you can do with vectors
```{r}
x <- c(1,2,3)

y <- c(100,200,300)

x + y

z <- x + y

z

```

R has tons of built in functions that you can apply to your data. Here are just a few examples..

```{r}
sum(3,4,2)

x <- c(3,4,2)

sum(x)

length(x)

mean(x)
```


You can put multiple vectors (columns of numbers) together to make a complete spreadsheet of data (a “data frame”). Most of the data we’ll work with will be organized in data frames.
```{r}
rat_number <- c(1,2,3)
weight <- c(8, 7, 12)

#You can store other things besides just numbers in variables in R.
rat_name <- factor(c("Joe", "Moe", "Doe"))

some_rat_weights <- data_frame(number = rat_number,
                              name = rat_name, weight = weight)

some_rat_weights 

#You can see your data frame in a really nice viewer by double clicking its name in the right side or by typing
View(some_rat_weights)
```

R can make graphs!
```{r}  
ggplot(data = some_rat_weights, aes(x = number, y = weight)) + 
  geom_point()

ggplot(data = some_rat_weights, aes(x = name, y = weight)) + 
  geom_point()
```

#You can pull out individual vectors (columns) from your data frame.
```{r}
some_rat_weights$number

some_rat_weights$weight
```

You can add new columns to your data frame simply by making up a new column name and storing something in it
```{r}
some_rat_weights$age <- c(12,14,17)

some_rat_weights

some_rat_weights$male <- factor(c("Y","Y","N"))

some_rat_weights
```

You can still do math on and apply functions to the vectors in a data frame (like you would on separate vectors).
```{r}
srw$age + 1

srw$weight / srw$age

mean(srw$age)
```

Just like a single variable, you can "overwrite" the data in a data frame column name simply by storing a new value in it
```{r}
some_rat_weights$male <- c(1,1,0)

some_rat_weights

srw <- some_rat_weights

some_rat_weights$age <- some_rat_weights$age + 1

some_rat_weights

srw <- some_rat_weights

srw
```


You can build a data frame from scratch (like above). OR you can import data from another source (for example, a .csv file)
```{r}
setwd("~/Desktop")

mnrd <- read_csv("Rat_Weights.csv")

mnrd

mnrd$Weight
```

FYI - in R studio you can run multiple lines of code at once by highlighting them and then clicking "Run" or "command + enter".

Other R studio features: Scroll back through past plots, erase all plots, erase all variables, save all varaibles for next session
