
---
title: "The Central Limit Theorem"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    toc: yes
editor_options:
  chunk_output_type: inline
---

```{r, message = F}
library(tidyverse)
theme_set(theme_classic(base_size = 16))
```

Read in the data
```{r read_csv, message = F}
# read class data
class_data <- read_csv("https://dyurovsky.github.io/psyc201/data/demos/psyc201_data.csv")

# read in Data from GSS 2016
sibs_data <- read_csv("https://dyurovsky.github.io/psyc201/data/demos/siblings.csv")
```

Number of siblings reported in the General Social Survey in 2016
```{r gss_sibs}
# plot gss siblings
ggplot(sibs_data, aes(x = siblings)) + 
  geom_histogram(binwidth = 1)  +
  scale_x_continuous(limits = range(sibs_data$siblings))
```

Number of siblings reported PsychStats in 2018
```{r class_sibs}
# plot class siblings
ggplot(class_data, aes(x = siblings)) + 
  geom_histogram(binwidth = 1)  +
  scale_x_continuous(limits = range(class_data$siblings))
```

Why isn't the class data normally distributed? Maybe it's not a representative sample? Let's take a sample

```{r sample_sibs} 
class_size_sample <- sibs_data %>%
  sample_n(nrow(class_data))


ggplot(class_size_sample, aes(x = siblings)) + 
  geom_histogram(binwidth = 1)  +
  scale_x_continuous(limits = range(class_size_sample$siblings))

```

Maybe that sample is too small?
```{r sample_sibs_bigger} 
big_sample <- sibs_data %>%
  sample_n(400)


ggplot(big_sample, aes(x = siblings)) + 
  geom_histogram(binwidth = 1)  +
  scale_x_continuous(limits = range(class_size_sample$siblings))

```

This can't be the problem, right? Because, as we keep increasing sample size, in the limit the sample becomes the population.

```{r sample_mean_function}
sample_mean <- function(size) {
  sibs_data %>%
    sample_n(size) %>%
    summarise(mean = mean(siblings)) %>%
    pull()
}
 

sample_mean(100)   
```


```{r distribution_of_sample_means_10}
sample_10 <- data_frame(sample = 1:1000, siblings = replicate(1000, sample_mean(10)))
  
ggplot(sample_10, aes(x = siblings)) + 
  geom_histogram(bins = 30)  +
  scale_x_continuous(limits = range(sample_10$siblings))
```

```{r distribution_of_sample_means_20}
sample_20 <- data_frame(sample = 1:1000, siblings = replicate(1000, sample_mean(20)))
  
ggplot(sample_20, aes(x = siblings)) + 
  geom_histogram(bins = 30)  +
  scale_x_continuous(limits = range(sample_20$siblings))
```


```{r distribution_of_sample_means_100}
sample_100 <- data_frame(sample = 1:1000, siblings = replicate(1000, sample_mean(100)))
  
ggplot(sample_100, aes(x = siblings)) + 
  geom_histogram(bins = 30)  +
  scale_x_continuous(limits = range(sample_100$siblings))
```

Let's see if the number of average siblings in the class is surprising given the GSS surver


```{r distribution_of_sample_means_nclass}
sample_nclass <- data_frame(sample = 1:1000, siblings = replicate(1000, sample_mean(nrow(class_data))))
  
ggplot(sample_nclass, aes(x = siblings)) + 
  geom_histogram(bins = 30)  +
  scale_x_continuous(limits = c(0, 6)) + 
  geom_vline(aes(xintercept = mean(class_data$siblings)), color = "darkred")
```